---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 10)
chart_2000 <- read.csv("data/chart2000-songyear-0-3-0058.csv",
                       stringsAsFactors = FALSE)
library(dplyr)
library(leaflet)
library(knitr)
library(ggplot2)
library(tidyr)

```

## Introduction

In this project, we will explore how music genres, artists, and songs have risen
and fallen in popularity over time.  We are interested in artist trends in the
U.S. because we want to see how artist preferences have changed over the
years, what artists have stayed relevant or popular, and what artists have
fallen out of popularity. Music is an integral part of society, and
culture and music taste can reflect societal attitudes, especially those
held by young people. Everyone in our group is an avid music listener, and we
are interested in gaining some insight into how music popularity has evolved
over time. In our analysis of this subject we worked with 2 seperate datasets.
The first dataset we explored was a compiled information on songs and artists
that gained enough popularity to be listed on the Billboard Hot 100 list from
2000 to 2020, which is a list of the most listened to songs in the U.S.A. We
used this dataset to find summary information concerning which songs were most
listened to and which artists made the most revenue, among other things. We
utilized the Spotify API to extract supplementary data for the songs in our
first dataset. We were able to use this data to search for popular artists that
showed up in the analysis of previous dataset, and discover what genre the song
or artist writes their music in. We also used this data in later charts to help
visualize the analysis we did, which you will see later in our report.


``` {r summary_list, echo = FALSE}

source("scripts/summary info function.R")
summary_list <- get_summary_info(chart_2000)
```

## Summary Information

To get an overview of this topic, we can look at an the data from a higher level
to make immediate inferences about our data set. 

Our team first looked at which artist showed up the most times on the Top 100
Chart in the last 20 years. We used this as a way to measure popularity. Based
on this measurement, we found that the artist who appeared the most times on the
chart was `r summary_list[5]`. This artist appeared a total of
`r summary_list[6]` times.

We also wanted to know which artist made the most total revenue in the last 20 
years to see what revenue says about how popular an artist is. As it turns out,
revenue and popularity go hand in hand. The artist with the most revenue was
again, `r summary_list[1]`, who made a total of $`r summary_list[2]` over the
last 20 years.

Switching gears from looking at artists, we took a look at trends pertaining to
songs. We found that the most popular songs on the charts were 
*`r summary_list[[9]][1]`* and *`r summary_list[[9]][2]`*. Popularity was
measured the same for songs as it was for artists.

Next, our team wanted to see which song made the most revenue in one year and 
which song made the least amount of revenue in one year. Just like with artist
revenue, song revenue shows whether a song was a real hit, or a real flop.
We found that the song that made the most amount of revenue in one year was
*`r summary_list[3]`*. This occured during the year `r summary_list[4]`.
The song that made the least amount of revenue in one year was
*`r summary_list[7]`*, which occured during the year `r summary_list[8]`.

## Summary Table

Our group was interested how the top artists have changed over the last 20
years. We decided to define "top artist" as the artist who reached position one
on the billboard hot 100 chart in each year. To explore this area of the data,
we aggregated data from our chart 2000 dataset to show which artists were in
position 1 each year. We grouped this data by year first to collect songs from
each individual year, then filtered for songs from each year that reached
position 1. The following data shows the arists from each year that reached
position one that we deemed as the top artist. We arranged this data in
descending order to see the most recent years first. 

```{r summary_table, echo=FALSE, warning=FALSE}
source("scripts/summary table function.R")
summary_table <- get_summary_table(chart_2000)

kable(summary_table, row.names = FALSE, col.names = c("Year","Top Artist"))
```

## Table insights

The first thing we notice from this table is that Ed Sheeran was the only
artist in the last 20 years that reached the top of the Billboard charts twice.
Not only did he reach the top twice, but he did it consecutively in both 2017
and 2018. We can also note who the most recent (2020) top artist is, Roddy
Ricch. We might be interested in the most recent top artist because their music
could be interpreted emblematic of the current most popular genre. Furthermore,
We can note the bottom-most artist in the table, which describes the top
arist in 2000, nearly two decades ago. This information serves as a description
of the state of music genere popularity in the past, and allows us to notice the
change in music style popularity over time. 

## Charts 

### Chart 1: Top Five Artists and 'Hits' Over Time

This chart shows how many hit songs the top five artists (Rihanna, Pink,
Maroon 5, The Black Eyed Peas, and Taylor Swift) from 2000 to 2020 had each
year. A 'hit' or 'hit song' is defined in this dataset and chart to be a song
that was included in the top 100 songs of that year, that being, the top 100 
songs that generated the most revenue. This chart may also serve as a way to
measure an artist's popularity throughout 2000 - 2020.

Pertinent observations from this chart include:

- The Black Eyed Peas had the highest amount of hits from 2000 - 2020 in 2010
when they had 5 hit songs.
- The Black eyed Peas stopped having any hit songs after 2011.
- Taylor Swift did not have any top hits until 2007.
- Rihanna's amount of hit songs in a given year appears to peak
from 2008 - 2011, where throughout 2008 and 2010 she consistently had 3 hit
songs each year and in 2011 she peaked at 4 hit songs.


```{r show chart 1, fig.width=12, fig.height=10, echo=FALSE}
source("scripts/Chart_1.R")
# Irrelevant column data is removed
stripped_df <- subset(chart_2000, select = -c(indicativerevenue, song, us, uk,
                                              de, fr, ca, au))

# Data frame of number of hits the top 5 aritsts from 2000-2020 had
filtered_df <- subset(stripped_df, artist %in% c("Rihanna", "Pink", "Maroon 5",
                                                 "The Black Eyed Peas",
                                                 "Taylor Swift")) %>%
  group_by(artist, year) %>%
  summarise(number_of_hits = n())

top5_num_hits <- get_top5_num_hits(filtered_df)
top5_num_hits
```

### Chart 2:
Chart 2 is a visualization of the relationship between genre popularity and
time. This was the best choice for showing how an individual genre's popularity
changed over time. To measure popularity, we took the top 10 artists from each
year and found what genre they write in. Then, we looked at how many times each
genre appeared during each year. There were handful of genres that only 
appeared once or twice over the 21 year period, so we ommited those results.

Pertinent observations from this chart include:
- Pop dominated the charts
- Pop was the only genre to stay relevent for the 21 year period
- Some genres started to become popular in the middle of the time span, like
rap, which became relevant in 2015 and stayed on the chart every year since

```{r show chart 2 + suppress warnings, message=FALSE, warning=FALSE, include=FALSE}
source("scripts/Chart_2.R")
non0_genres <- chart %>% 
  group_by(year, genre) %>% 
  summarize(times = n())

with_zeros <- non0_genres %>% 
  spread(genre, times, fill=0) %>% 
  gather(genre, times, -year) %>%
  filter(genre != "country",
         genre != "big room", 
         genre != "boy band",
         genre != "crunk",
         genre != "complextro",
         genre != "dance",
         genre != "electro",
         genre != "funk",
         genre != "irish singer-songwriter",
         genre != "permanent wave")
plot <- get_top_10_genres(with_zeros)
```
``` {r show plot, echo = FALSE}
plot
```
### Chart 3:

Chart 3 is a visual discription of the total indicative revenue of the top songs
of 2012 to 2020. Indicative revenue is the complete revenure generated by a song
or album over a certain periond. It takes inflation and currency conversion into
account. In the chart there are a total of nine songs that are being tracked. 
These songs are Somebody That I Used To Know, Blurred Lines, Happy, Uptown Funk,
Cheap Thrills, Shape of You, Perfect, Old Town Road, and The Box. 

Pertinent observations from this chart include:
- In 2017 Shape of you made the most compared to all the other top songs of
their year
- With Old Town Road coming in second in the year of 2019 
- Unlike the Shape of You, The Box made the least compare to the rest of the
songs of their year. 

``` {r show chart 3, echo=FALSE}
source("scripts/Chart_3.R")
# Data frame of the songs, position and revenue 
song_revenue_position_df <- subset(chart_2000, select = -c(artist, us, uk,
                                             de, fr, ca, au))
# Top songs of each year 
topsongs_year <- song_revenue_position_df %>% 
  filter(position == 1) %>% 
  filter(year > 2011) %>% 
  subset(select = -c(position))

plot_topsongs <- get_money_for_topsongs(topsongs_year)
plot_topsongs
```

